Factor Analysis on Citations, Using a Combined Latent and Logistic Regression Model

### 2019-10-07 Manuscript for submission/manuscript.tex
Update

\subsubsection{Synthetic Setting} \label{sec:Syn-set}
%In this subsection, we consider the following $3$ scenarios:
%In this subsection, we consider $3$ scenarios.
%In first scenario, we examine a simple setting, in which every paper in the network has only one topic.
%In second scenario, we study a more complicated situation than the first one where each paper in the network can possibly have more than one topic.
%It is worth noting that different clusters of papers, where each cluster is comprised of papers with same topic, are connected via ad-hoc links.
%In third scenario, we consider a situation
Before specifying the network settings in two scenarios that we want to explore, let us first describe a set of procedures for setting the model parameters, $\alpha^*, F^*, D^*$ and $S^*$ sequentially.

@@

    \item Recall that the binary factor loading matrix $F^*$ encodes the relation between topics and papers
        (i.e., if $i$th paper studies about $k$th topic, we denote $F^*_{ki}=1$, otherwise $F^*_{ki}=0$).
        First, we assume that there are $n$ papers in the network, and $K$ topics are embedded in it.
        Each of them consists of $\frac{n}{K}$ papers.
        This can be expressed in $F^*$ as follows:
        \[
            F^* =
               \begin{bmatrix}
                 \smash[b]{\block{n/K}} \\
                 && \smash[b]{\block{n/K}} \\
                 &&&& \ddots \\
                 &&&&& \block{n/K}
                \end{bmatrix}
              \in R^{K \times n},
        \]
        where each row of $F^*$ has $\frac{n}{K}$ $1$'s and each column has only one $1$.
        Then, we assume that $n_{l}$ papers share $l$ topics and $n_{m}$ papers share $m$ topics ($1 < l < m \leq K$).
        This can be represented in the $F^*$ in following steps:
        $1)$ Pick distinct $n_{l}$ indices randomly from $\{1,2,\dots,n\}$.
        We will denote the set of the indices as $\Omega_{n_l}$.
        $2)$ Choose $n_{m}$ indices from the set $\{1,2,\dots,n\}\setminus \Omega_{n_{l}}$ and denote the set of those indices as $\Omega_{n_{m}}$.
        $3)$ Make the columns of $F^*$ with corresponding indices in set $\Omega_{n_{l}} \cup \Omega_{n_{m}}$ zeros.
        $4)$ We use a notation $f^*_j$ to denote the $j$th column of the matrix $F^*$.
        Fill arbitrary $l$ entries of $f^*_j$ for $j \in \Omega_{n_l}$ with $1$s, and also fill arbitrary $m$ entries of $f^*_j$ for $j \in \Omega_{n_m}$ with $1$s.
        Lastly, we set $F^*=JF^*$ where $J=I_{n}-\frac{1}{n}\mathbbm{1}\mathbbm{1}^T$.

@@

\begin{figure}[htbp]
\includegraphics[width=1\textwidth]{Fig1.pdf}
\caption{Graphical illustrations of six synthetic networks.
Nodes that share the common factors are clustered. The cross cluster links are ad-hoc Citations.
All the graphs are drawn via R package `igraph'.}
\label{fig:figure1}
\end{figure}

@@

Non-zero components of $\widehat{S}$ capture the citation relationships among papers that are not attributable to the common topics.
The selected model has $51$ sparse edges, and all of them are positive edges.
In Table \ref{tab:table2}, we provide $10$ pairs of papers that have the largest estimated $\widehat{S}_{ij}$.
All the $10$ edges come from the pairs of papers from different communities.
For instance, the first pair of papers comes from the Functional Analysis community and Variable Selection community.
$\emph{FuncAn}$ paper cites $\emph{VarSel}$ paper for borrowing a mathematical representation to build a theorem.
Though it might appear to be a crucial step for building a theorem in their paper, we cannot say that two papers are closely related in terms of topic.
The second pair of papers comes from the Mixed topics community (denoted as $\emph{Mixed}$) and the Variable Selection ($\emph{VarSel}$) community.
This case is interesting since both papers study about variable selection problem, but they are classified in different communities and connected via an ad-hoc link.
Specifically, the paper from $Mixed$ community study the variable selection problem under the non-parametric Bayesian framework, and compare their method with the one that is proposed in $\emph{VarSel}$ paper, ``Adaptive Lasso’’.

>>>>

\subsubsection{Synthetic Setting} \label{sec:Syn-set}
%In this subsection, we consider the following $3$ scenarios:
%In this subsection, we consider $3$ scenarios.
%In first scenario, we examine a simple setting, in which every paper in the network has only one topic.
%In second scenario, we study a more complicated situation than the first one where each paper in the network can possibly have more than one topic.
%It is worth noting that different clusters of papers, where each cluster is comprised of papers with same topic, are connected via ad-hoc links.
%In third scenario, we consider a situation
Before specifying the network settings in two scenarios that we want to explore, let us first describe a set of procedures for setting the model parameters, $\alpha^*, F^*, D^*$ and $S^*$ sequentially.
We put stars to the superscript of parameters to denote that they are ground truth.
Readers can refer the meaning of each parameter in the model in Section \ref{sec:intro} and Section \ref{sec:model-form}.

@@

    \item Recall that the binary factor loading matrix $F^*$ encodes the relation between topics and papers
        (i.e., if $i$th paper studies about $k$th topic, we denote $F^*_{ki}=1$, otherwise $F^*_{ki}=0$).
        First, we assume that there are $n$ papers in the network, and $K$ topics are embedded in it.
        Each of them consists of $\frac{n}{K}$ papers.
        This can be expressed in $F^*$ as follows:
        \[
            F^* =
               \begin{bmatrix}
                 \smash[b]{\block{n/K}} \\
                 && \smash[b]{\block{n/K}} \\
                 &&&& \ddots \\
                 &&&&& \block{n/K}
                \end{bmatrix}
              \in R^{K \times n},
        \]
        where each row of $F^*$ has $\frac{n}{K}$ $1$'s and each column has only one $1$.
        Note that the remaining entries of the matrix are filled with zeros.
        Then, we assume that $n_{l}$ papers share $l$ topics and $n_{m}$ papers share $m$ topics ($1 < l < m \leq K$).
        This can be represented in the $F^*$ in following steps:
        $1)$ Pick distinct $n_{l}$ indices randomly from $\{1,2,\dots,n\}$.
        We will denote the set of the indices as $\Omega_{n_l}$.
        $2)$ Choose $n_{m}$ indices from the set $\{1,2,\dots,n\}\setminus \Omega_{n_{l}}$ and denote the set of those indices as $\Omega_{n_{m}}$.
        $3)$ Make the columns of $F^*$ with corresponding indices in set $\Omega_{n_{l}} \cup \Omega_{n_{m}}$ zeros.
        $4)$ We use a notation $f^*_j$ to denote the $j$th column of the matrix $F^*$.
        Fill arbitrary $l$ entries of $f^*_j$ for $j \in \Omega_{n_l}$ with $1$s, and also fill arbitrary $m$ entries of $f^*_j$ for $j \in \Omega_{n_m}$ with $1$s.
        Lastly, we set $F^*=JF^*$ where $J=I_{n}-\frac{1}{n}\mathbbm{1}\mathbbm{1}^T$.

@@

\begin{figure}[htbp]
\includegraphics[width=1\textwidth]{Fig1.pdf}
\caption{Graphical illustrations of six synthetic networks.
Nodes that share the common factors are clustered. The cross cluster links are ad-hoc Citations.
All the graphs are drawn via the algorithm in \cite{fruchterman1991graph}.}
\label{fig:figure1}
\end{figure}

@@

Non-zero components of $\widehat{S}$ capture the citation relationships among papers that are not attributable to the common topics.
The selected model has $51$ sparse edges, and all of them are positive edges.
In Table \ref{tab:table2}, we provide $10$ pairs of papers that have the largest estimated $\widehat{S}_{ij}$.
All the $10$ edges come from the pairs of papers from different topics.
For instance, the first pair of papers comes from the Functional Analysis topic and Variable Selection topic.
The paper from Functional Analysis topic cites the paper from Variable Selection for borrowing a mathematical representation to build a theorem.
Though it might appear to be a crucial step for building a theorem in their paper, we cannot say that two papers are closely related in terms of topic.
The second pair of papers comes from the Mixed topics community and the Variable Selection community.
This case is interesting since both papers study about variable selection problem, but they are classified in different communities and connected via an ad-hoc link.
Specifically, the authors in the paper from Mixed topics community study the variable selection problem under the non-parametric Bayesian framework, and compare their method with the ``Adaptive Lasso'' that is proposed in the paper from Variable Selection topic.
